Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal density\thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-11-21 manuscript_v2_noiseSim.tex
edit Xav

\documentclass{aa}
%\documentclass[onecolumn]{aa}
%\documentclass[referee]{aa}
\usepackage{pdflscape} % or {lscape}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
%\usepackage{hyperref}
\usepackage[draft]{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}

\usepackage{afterpage}
\usepackage{ulem}

@@

%{The performances of the proposed method are compared to the commonly employed Normal density using both simulations and real observations, with different levels of activity and signal-to-noise ratio (S/N) levels.}

@@

{The performances of the proposed method are compared to the commonly employed Normal density using both simulations and real observations, with different levels of activity and signal-to-noise ratio (S/N) levels.}

@@

when estimating stellar rotational periods and more in general when characterizing stellar activity signals.

@@

compared to the usual model used to account for stellar activity.

@@

One of the most challenging stellar signals to characterize and to correct for are the signals induced by stellar activity. 
Stellar activity is responsible for creating magnetic regions on the surface of stars, and those regions change locally the temperature and the convection, which can induce spurious RVs variations \citep[e.g.][]{Meunier-2010a, Dumusque-2014b, Borgniet-2015}. 
In theory, it should be easy to differentiate between the pure Doppler-shift induced by a planet, which shifts the entire stellar spectrum, and stellar activity, which modifies the shape of spectral lines and by doing so creates a spurious shift of the stellar spectrum \citep{Saar-1997b,Hatzes-2002,Kurster2003,Lindegren-2003,Desort-2007,Lagrange-2010,Meunier-2010a,Dumusque-2014b}. 
However, on quiet GKM dwarfs, the main targets for precise RVs measurements, stellar activity can induce signals of a few \ms. 
This corresponds physically to variations smaller than 1/100th of a pixel on the detector, making changes in the shape of the spectral lines challenging to detect.
In order to measure such tiny variations, a common approach is to average the information of all the lines in the spectrum by cross correlating the stellar spectrum with a synthetic or an observed stellar template \citep[][]{Baranne-1996,Pepe-2002a, Anglada-Escude-2012}. The result of this operation gives us the cross-correlation function (CCF).  
%The CCF gives the spectrum's cross-correlation with the template as the template is shifted according to different RVs.
%
To measure the Doppler-shift between different spectra, and therefore to retrieve the RVs of a star as a function of time, the variations of the CCF barycenter are calculated. 
The barycenter is generally estimated by the mean of  a Normal density shape fit to the CCF. 
Variations in line shape between different spectra, which indicate the presence of signals induced by stellar activity, are measured by analyzing different parameters of models fit to the CCF. Usually the width of the CCF is estimated using the full-width half-maximum (FWHM) of the fitted Normal density and its asymmetry by calculating the CCF bisector and measuring the bisector inverse slope span \citep[BIS SPAN,][]{Queloz-2001}.

@@

To evaluate the strength of the correlation between the estimated RVs and the different stellar activity indicators, we calculated the Pearson correlation coefficient, $R$, which in its general form is defined as:
%
\begin{equation}
R (x,y)= \frac{\text{cov}(x,y)}{\sigma(x)\sigma(y)},
\label{eq:Pearson:corr}
\end{equation}
%
where $x$ and $y$ are two quantitative variables, $\text{cov}(x,y)$ indicates the covariance between $x$ and $y$, and $\sigma(x)$ and $\sigma(y)$ represent their standard deviations.  A $p$-value for the statistical test with null hypothesis $H_{0}: R= 0$ is also generally provided.

@@

\textbf{Exoplanets produce only a pure RV signal (i.e., exoplanets induce a Doppler shift of the entire stellar spectrum)}. On the contrary stellar activity and, in particular, the presence of active regions on the stellar photosphere, do not produce blueshifts or redshifts of the entire stellar spectrum but can create spurious RV signals by modifying the shape of spectral lines.
To track these variations in the shape of the spectral lines, a common approach consists in using the FWHM, the BIS SPAN, or other indicators such as those introduced in \citet{Boisse-2011} or \citet{Figueira-2013}, which provide information on the width and asymmetry of the CCF. A strong correlation between the estimated RVs and one or more of these parameters provides an indication that stellar activity signals may be affecting the measurements.

@@

For a given configuration of spots and faculae on the stellar surface, SOAP 2.0 outputs simulated CCFs as a function of rotational phase. 
The code also returns the RV and the FWHM estimated using a Normal model for the CCF, and the BIS SPAN obtained by calculating the bisector of the CCF. 
SOAP 2.0 gives noiseless CCFs affected by stellar activity, which are used to compare the performances of SN and Normal models of the CCF.

For the simulations discussed below, a star similar to the Sun was modelled with a solar disc of one solar radius seen equator-on, and a stellar rotational period set to 25.0 days.
The stellar effective temperature is set to 5778 K, and a quadratic limb-darkening relation with linear and quadratic coefficients 0.29 and 0.34, respectively \citep[][]{Oshagh-2013a, Claret-2011}. {\bf The temperature difference of the spot with the photosphere is $\Delta T_{spot} = -663$ K and the temperature difference of the facula depends on the centre-to-limb angle $\theta$,  $\Delta T_{plage} = 250.9 - 407.7\,\cos{\theta} + 190.9\,\cos^2 \theta$ K \citep[][]{Meunier-2010a}}.
In order to make the result of the simulations more comparable to real data obtained with the HARPS spectrograph discussed in Sect.~\ref{sec:4}, the SOAP 2.0 CCFs were generated with a width of 40 \kms\, and considering initial spectra with resolution of R= 115'000.
{\bf{ We consider an SNR= 100 for our simulations.}}

@@

Correlations between the different RV estimates and the different CCF asymmetry or width estimates are displayed in Fig.~\ref{fig:faculae.corr}. 
The correlation between $\gamma$ and SN mean RV, and $\gamma$ and SN median RV are {\bf{weaker}} than the correlation between BIS SPAN and RV, with Pearson correlation coefficient values of {\bf{$R$=-0.11, -0.55 and -0.81}}, respectively. 
%
{\bf{There is a stronger correlation between SN FWHM and SN mean RV ($R=0.92$) than between SN FWHM and SN median RV ($R=0.65$).  The correlation between FWHM and N mean RV is the weakest ($R=0.47$), and the correlation between FWHM and SN mean RV is the strongest ($R=0.92$).}} 
This first analysis shows that in the case of a facula, using some parameters from the SN can lead to stronger correlations than the ones obtained by the usual Normal parameters. 
{\bf Therefore, the SN parameters may better probe stellar activity that induces changes in the FWHM.}
We investigate this feature further in the following two sections, where we consider simulated data with a single spot and a spot plus a planet, and in Sec.~\ref{sec:4} with real observations.

@@

The RV variations displayed in Fig.~\ref{fig:faculae} are caused only by stellar activity, in this case a facula, so we applied the activity correction proposed in Eq.~\eqref{eq:RV:correction} to evaluate the performance of the model in this setting.
The results of this correction are displayed in Fig.~\ref{fig:faculae.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table \ref{table:faculae.test}. 
The proposed correction for stellar activity is able to account for the majority of the activity signal created by a facula, with a {\bf{$R^2$ larger than $0.98$.}}
%In addition, the rms of the different estimates of the RV reduces from about {\bf{6 \ms}} before correcting for stellar activity to values below {\bf{1.2 \ms}} after correcting for stellar activity.
%We see a slightly smaller rms after correction when using the SN parameters compared to the Normal parameters.
{\bf{In addition the rms for N mean RV reduces from 8.02 \ms before correcting for stellar activity to 1.02 \ms after correcting for stellar activity using the proposed linear model; the respective values for the SN median RV are 7.07 \ms to 0.88 \ms and for the SN mean RV are 5.9 \ms to 0.88 \ms.}}
When comparing the correction proposed in Eq.~\eqref{eq:RV:correction} with what is generally used (i.e. a linear combination of only the asymmetry and width parameters), we see that the proposed correction is able to reduce the rms of the RV residuals by {\bf{an additional 14.5 - 15\%}}. Looking at the significance of the coefficients in Table~\ref{table:faculae.test}, we observe that {\bf{all the parameters corresponding to the SN variables are statistically significant. When using the Normal variables, the parameters $\beta_{3}$ and $\beta_{4}$ are not statistically helpful for the correction.}}

\begin{figure*}[htbp]
\begin{center}
%\includegraphics[height = 6in]{FACULAE_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
\includegraphics[height = 6in]{FACULA_COMPARISON_withNoise.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a facula in the simulated data using a Normal and a SN model, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red pluses), and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the Normal fit.    
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red pluses) and the residuals from the usual correction (green triangles). The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model of Eq.~\eqref{eq:RV:correction} compared to the usual model.
The tests of statistical significance on the parameters are presented in Table \ref{table:faculae.test}.
}\label{fig:faculae.correction}
\end{center}
\end{figure*}

@@

Fig.~\ref{fig:spot.corr} shows the correlations between the asymmetry parameters and the different estimates of the CCF barycenter (i.e. SN mean RV, SN median RV and N mean RV). {\bf{The correlation between $\gamma$ and SN median RV and the correlation between the BIS SPAN and N mean RV are the strongest ($R=-0.81$) followed by the $\gamma$ - SN mean RV correlation ($R=-0.76$).}} 
The correlations between the width and the CCF barycenter are more circular and no substantial correlation is observed. 
%Similar to the case with a facula, some parameters of the SN gives stronger correlations compared to the Normal parameters. {\color{red} \bf I'm not sure the previous statement is correct anymore - the only correlations strong with the SN variables is described has having ``no substantial correlation'' in the previous statement...I suggest we delete the previous sentence, or revise to make it correct.]}

@@

As before, the original RV estimates are corrected using Eq.~\eqref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
%{\bf{In this case the proposed correction is not able to perform as well as for the facula, and $R^2$ values for the linear combination are smaller than 0.8}}. 
{\bf The $R^2$ values for the linear combination are between 0.72 and 0.78, which are not as strong as the setting with a facula.}

%In Fig.~\ref{fig:spot.correction}, we see that the proposed activity correction is able to reduce the signal of a spot from a raw RV rms larger than {\bf{6.14\ms}} down to a rms smaller than {\bf{3.05 \ms}}, for any of the different RV estimates.
{\bf{In addition the rms for N mean RV reduces from 6.14 \ms before correcting for stellar activity to 3.04 \ms after  correcting for stellar activity using the proposed linear model; the respective values for the SN median RV are 5.85 \ms to 2.74 \ms and for the SN mean RV are 5.27 \ms to 2.74 \ms.}}
When comparing the activity correction proposed in this paper with what is commonly used, i.e only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by {\bf an additional 5.3 - 5.8\%}.
The Normal or SN parameters involved in Eq.~\eqref{eq:RV:correction} are all statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, {\bf{except the width of the CCF and the variable that captures the interaction between the width and the asymmetry of the CCF.}}
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

\begin{figure*}[htbp]
\begin{center}
%\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf}
\includegraphics[height = 6in]{SPOT_COMPARISON_withNoise.pdf}
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red pluses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red pluses) and the residuals from the usual correction (green triangles). 
 The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model compared to the usual model.
The tests of statistical significance on the parameters are presented in Table~\ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

@@

The correlation between the different CCF parameters are displayed in Fig.~\ref{fig:spot.plus.planet.corr}. 
The correlations are weaker than in the case of the spot due to the planet inducing changes in RV without affecting the width or the asymmetry of the CCF.  However, the order of the strength of the correlations between the CCF asymmetry parameters and RV are comparable with the ones obtained for the spot-only model: $\gamma$--SN median RV has the strongest correlation {\bf{(R= -0.7)}}, followed by the correlation between BIS SPAN--N mean RV {\bf{(R= -0.69)}} and finally by the correlation between $\gamma$--SN mean RV {\bf{(R= -0.67)}}. 
The patterns observed in the width-RV phase space correlations in Fig.~\ref{fig:spot.plus.planet.corr} follow a circle, similar to the spot-only model; no substantial correlation is observed between those two parameters.

@@

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\eqref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\eqref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\eqref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table~\ref{table:spotplanet.test}.   
{\bf{Except for the intercept term (coefficient $\beta_0$), the width of the CCF (coefficient $\beta_3$), and the variable that captures the interaction between the width and the asymmetry of the CCF (coefficient $\beta_4$), all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. 
We also observe that the RV residuals, after correcting for stellar activity and the presence of the planet, have smaller rms when using the SN variables (rms = 2.29 \ms) rather than the Normal ones (rms = 2.80 \ms).}}
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\eqref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\eqref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the different coefficients used in Eq.~\eqref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with a period of one-third the rotational period of the star and a semi-amplitude  of 10\,\ms. {\bf{All the parameters corresponding to the Normal or SN variables are statistically significant, except the intercept, the width of the CCF and the variable that evaluates the interaction between the width and the asymmetry of the CCF. Once corrected for stellar activity and the presence of the planet, the residuals are smaller in terms of rms when using the SN variables}}. Note that since nonlinear least squares was required, the residual standard error rather than the $R^2$ is displayed as a reference.}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.95$    & $0.076$  & $0.21$ \\
\hline
$\beta_{1}$            &    $0.0020$    & $0.11e-4$  & $0.12e-4$ \\
\hline
$\beta_{2}$            &     $2.22e-16$   & $2.22e-16$ & $2.22e-16$\\
\hline
$\beta_{3}$            &     $0.92$   &  $0.60$  & $0.61$\\
\hline
$\beta_{4}$            &     $0.34$   &  $0.27$ & $0.27$\\
\hline
K            &     $2.22e-16$   &  $2.22e-16$   & $2.22e-16$ \\
\hline
P            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$ \\
\hline
$t_{0}$            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$ \\
\hline
$\text{Residuals}$      &     $2.80 \ms$    &  $ 2.29 \ms$ & $2.29 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

\documentclass{aa}
%\documentclass[onecolumn]{aa}
%\documentclass[referee]{aa}
\usepackage{pdflscape} % or {lscape}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
%\usepackage{hyperref}
\usepackage[draft]{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}
\usepackage[normalem]{ulem}

\usepackage{afterpage}

@@

%{The performances of the proposed method are compared to the commonly employed Normal density using both simulations and real observations, with different levels of activity and signal-to-noise ratio.}

@@

{The performances of the proposed method are compared to the commonly employed Normal density using both simulations and real observations, with different levels of activity and signal-to-noise ratio.}

@@

when estimating stellar rotational periods and when characterizing stellar activity signals.

@@

compared to the model commonly used to account for stellar activity.

@@

One of the most challenging stellar signals to characterize and to correct for are the signals induced by stellar activity. 
Stellar activity is responsible for creating magnetic regions on the surface of stars, and those regions change locally the temperature and the convection, which can induce spurious RVs variations \citep[e.g.][]{Meunier-2010a, Dumusque-2014b, Borgniet-2015}. 
In theory, it should be easy to differentiate between the Doppler-shift induced by a planet, which shifts the entire stellar spectrum, and the effect of stellar activity, which modifies the shape of spectral lines and by doing so creates a spurious shift of the stellar spectrum \citep{Saar-1997b,Hatzes-2002,Kurster2003,Lindegren-2003,Desort-2007,Lagrange-2010,Meunier-2010a,Dumusque-2014b}. 
However, on quiet GKM dwarfs, the main targets for precise RVs measurements, stellar activity can induce signals of a few \ms. 
This corresponds physically to variations smaller than 1/100th of a pixel on the detector, making changes in the shape of the spectral lines challenging to detect.
In order to measure such tiny variations, a common approach is to average the information of all the lines in the spectrum by cross correlating the stellar spectrum with a synthetic or an observed stellar template \citep[][]{Baranne-1996,Pepe-2002a, Anglada-Escude-2012}. The result of this operation gives us the cross-correlation function (CCF).  
%The CCF gives the spectrum's cross-correlation with the template as the template is shifted according to different RVs.
%
To measure the Doppler-shift between different spectra, and therefore to retrieve the RVs of a star as a function of time, the variations of the CCF barycenter are calculated. 
The barycenter is generally estimated by the mean of  a Normal density shape fit to the CCF. 
Variations in line shape between different spectra, which indicate the presence of signals induced by stellar activity, are measured by analyzing different parameters of models fit to the CCF. Usually the width of the CCF is estimated using the full-width half-maximum (FWHM) of the fitted Normal density and its asymmetry by calculating the CCF bisector and measuring the bisector inverse slope span \citep[BIS SPAN,][]{Queloz-2001}.

@@

To evaluate the strength of the correlation between the estimated RVs and the different stellar activity indicators, we calculated the Pearson correlation coefficient, $R$, which in its general form is defined as:
%
\begin{equation}
R (x,y)= \frac{\text{cov}(x,y)}{\sigma(x)\sigma(y)},
\label{eq:Pearson:corr}
\end{equation}
%
where $x$ and $y$ are two quantitative variables, $\text{cov}(x,y)$ indicates the covariance between $x$ and $y$, and $\sigma(x)$ and $\sigma(y)$ represent their standard deviations.  A $p$-value for the statistical test with null hypothesis $H_{0}: R=0$ is also generally provided.

@@

\textbf{Exoplanets produce a Doppler shift of the entire stellar spectrum}. On the contrary stellar activity and, in particular, the presence of active regions on the stellar photosphere, do not produce blueshifts or redshifts of the entire stellar spectrum but can create spurious RV signals by modifying the shape of spectral lines.
To track these variations in the shape of the spectral lines, a common approach consists in using the FWHM, the BIS SPAN, or other indicators such as those introduced in \citet{Boisse-2011} or \citet{Figueira-2013}, which provide information on the width and asymmetry of the CCF. A strong correlation between the estimated RVs and one or more of these parameters provides an indication that stellar activity signals may be affecting the measurements.

@@

For a given configuration of spots and faculae on the stellar surface, SOAP 2.0 outputs simulated CCFs as a function of rotational phase. 
The code also returns the RV and the FWHM estimated using a Normal model for the CCF, and the BIS SPAN obtained by calculating the bisector of the CCF. 
\sout{SOAP 2.0 gives noiseless CCFs affected by stellar activity, which are used to compare the performances of SN and Normal models of the CCF.}

For the simulations discussed below, a star similar to the Sun was modelled with a solar disc of one solar radius seen equator-on, and a stellar rotational period set to 25.0 days.
The stellar effective temperature is set to 5778 K, and a quadratic limb-darkening relation with linear and quadratic coefficients 0.29 and 0.34, respectively \citep[][]{Oshagh-2013a, Claret-2011}. {\bf The temperature difference of the spot with the photosphere is $\Delta T_{spot} = -663$ K and the temperature difference of the facula depends on the centre-to-limb angle $\theta$,  $\Delta T_{facula} = 250.9 - 407.7\,\cos{\theta} + 190.9\,\cos^2 \theta$ K \citep[][]{Meunier-2010a}}.
In order to make the result of the simulations more comparable to real data obtained with the HARPS spectrograph discussed in Sect.~\ref{sec:4}, the SOAP 2.0 CCFs were generated with a width of 40 \kms\, and considering initial spectra with resolution of R=115'000.
{\bf{ For the simulations presented in this section, we considered a S/N of 100.}}

@@

Correlations between the different RV estimates and the different CCF asymmetry or width estimates are displayed in Fig.~\ref{fig:faculae.corr}. 
{\bf The correlation between $\gamma$ and SN mean RV, and $\gamma$ and SN median RV are weaker than the correlation between BIS SPAN and RV, with Pearson correlation coefficient values of $R$=-0.11, -0.55 and -0.81, respectively. Therefore, it seems that when looking at the CCF asymmetry, fitting a SN density to the CCF does not help. However, when looking at the correlations between the width of the CCF and the RV, the stronger correlation can be found between SN FWHM and SN mean RV ($R=0.92$). Then follows the correlation between SN FWHM and SN median RV ($R=0.65$), and finally FWHM and N mean RV ($R=0.47$). Over all the different correlations in the case of a facule, the strongest one is found between SN FWHM and SN mean RV, which demonstrates that fitting a SN density to the CCF can be helpful to better probe stellar activity}
%%
%{\bf{There is a stronger correlation between SN FWHM and SN mean RV ($R=0.92$) than between SN FWHM and SN median RV ($R=0.65$), but the correlation between FWHM and N mean RV is the weakest ($R=0.47$).}} 
%This first analysis shows that in the case of a facula, using some parameters from the SN can lead to stronger correlations than the ones obtained by the usual Normal parameters. Therefore, the SN parameters may better probe stellar activity. 
%We investigate this feature further in the following two sections, where we consider simulated data with a single spot and a spot plus a planet, and in Sec.~\ref{sec:4} with real observations.

@@

The RV variations displayed in Fig.~\ref{fig:faculae} are caused only by stellar activity, in this case a facula. We applied the activity correction proposed in Eq.~\eqref{eq:RV:correction} to evaluate the performance of the model in this setting.
The results of this correction are displayed in Fig.~\ref{fig:faculae.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table \ref{table:faculae.test}. 
The proposed correction for stellar activity is able to account for the majority of the activity signal created by a facula, with a {\bf{$R^2$ larger than $0.98$.}}
%In addition, the rms of the different estimates of the RV reduces from about {\bf{6 \ms}} before correcting for stellar activity to values below {\bf{1.2 \ms}} after correcting for stellar activity.
%We see a slightly smaller rms after correction when using the SN parameters compared to the Normal parameters.
{\bf In addition, the proposed linear model allows to reduce the activity effect from a RV rms of 8.02 \ms to 1.02 \ms when considering the CCF contrast and FWHM, and the BIS SPAN. When using the parameters derived from the SN, the improvements for SN median RV and SN mean RV are 7.07 \ms down to 0.88 \ms and 5.9 \ms down to 0.88 \ms, respectively}
When comparing the correction proposed in Eq.~\eqref{eq:RV:correction} with what is generally used (i.e. a linear combination of only the asymmetry and width parameters), we see that the proposed correction is able to reduce the rms of the RV residuals by  {\bf{15 \%}}. Looking at the significance of the coefficients in Table~\ref{table:faculae.test}, we observe that {\bf{all the parameters corresponding to the SN variables are statistically significant. When using the Normal variables, the parameters $\beta_{3}$ and $\beta_{4}$ are not statistically helpful for the correction.}}

\begin{figure*}[htbp]
\begin{center}
%\includegraphics[height = 6in]{FACULAE_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
\includegraphics[height = 6in]{FACULA_COMPARISON_withNoise.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a facula in the simulated data using a Normal and a SN model, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red pluses), and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the Normal fit.    
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red pluses) and the residuals from the usual correction (green triangles). The standard deviations are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model of Eq.~\eqref{eq:RV:correction} compared to the usual model.
The tests of statistical significance on the parameters are presented in Table \ref{table:faculae.test}.
}\label{fig:faculae.correction}
\end{center}
\end{figure*}

@@

Fig.~\ref{fig:spot.corr} shows the correlations between the asymmetry parameters and the different estimates of the CCF barycenter (i.e. SN mean RV, SN median RV and N mean RV). {\bf{The correlation between $\gamma$ and SN median RV and the correlation between the BIS SPAN and N mean RV are the strongest ($R=-0.81$) followed by the $\gamma$ - SN mean RV correlation ($R=-0.76$).}} 
The correlations between the width and the CCF barycenter draw circles and no significant correlation is observed. 
{\bf Not like in the facula case, when considering a spot simulated with SAOP 2.0 and a S/N of 100, the SN fit to the CCF does not gives parameters that can probe better stellar activity.
\sout{Similar to the case with a facula, some parameters of the SN gives stronger correlations compared to the Normal parameters.}}

@@

As before, the original RV estimates are corrected using Eq.~\eqref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
{\bf In the case of a spot, the proposed correction is not able to perform as well as for the facula, and $R^2$ values for the linear combination are between 0.7 and 0.8}. 
{\bf The correction is able to mitigate stellar activity from a N mean RV rms of 6.14\ms down to 3.04\ms. The corresponding values for SN median RV and SN mean RV are 5.85\ms down to 2.74 \ms and 5.27\ms down to 2.74\ms, respectively.}
When comparing the activity correction proposed in this paper with what is commonly used, which means only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by {\bf{6 \%}}.
The Normal or SN parameters involved in Eq.~\eqref{eq:RV:correction} are statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, {\bf except the width of the CCF $\beta_3$ and the interaction term $\beta_4$}
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

\begin{figure*}[htbp]
\begin{center}
%\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf}
\includegraphics[height = 6in]{SPOT_COMPARISON_withNoise.pdf}
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red pluses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red pluses) and the residuals from the usual correction (green triangles). 
 The standard deviations are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model compared to the usual model.
The tests of statistical significance on the parameters are presented in Table~\ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

@@

The correlation between the different CCF parameters are displayed in Fig.~\ref{fig:spot.plus.planet.corr}. 
The correlations are weaker than in the case of the spot due to the planet inducing changes in RV without affecting the width or the asymmetry of the CCF.  However, the order of the strength of the correlations between the CCF asymmetry parameters and RV are comparable with the ones obtained for the spot-only model: $\gamma$--SN median RV has the strongest correlation {\bf{($R=-0.7$)}}, followed by the correlation between BIS SPAN--N mean RV {\bf{($R=-0.69$)}} and finally by the correlation between $\gamma$--SN mean RV {\bf{($R=-0.67$)}}. 
The patterns observed in the width-RV phase space correlations in Fig.~\ref{fig:spot.plus.planet.corr} follow a circle, similar to the spot-only model; no substantial correlation is observed between those two parameters.

@@

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\eqref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\eqref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\eqref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table~\ref{table:spotplanet.test}.   {\bf{Except for the intercept $\beta_0$, the width of the CCF $\beta_3$, and the interaction term $\beta_4$ that evaluates the interaction between the width and the asymmetry of the CCF, all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. We also observe that the RV residuals, once corrected for stellar activity and the presence of the planet, are smaller in terms of rms when using the SN variables (rms = 2.28 \ms) rather than the Normal ones (rms = 2.80 \ms).}}
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\eqref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\eqref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the different coefficients used in Eq.~\eqref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with a period of one-third the rotational period of the star and a semi-amplitude  of 10\,\ms. {\bf All the parameters corresponding to the Normal or SN variables are statistically significant, except the intercept $\beta_0$, the width of the CCF $\beta_3$ and the interaction term $\beta_4$ that evaluates the interaction between the width and the asymmetry of the CCF. Once corrected for stellar activity and the presence of the planet, the residuals are smaller in terms of rms when using the SN variables. Note that because we used nonlinear least squares to fit the best model, the residual rms rather than the $R^2$ is displayed as a reference.}}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.95$    & $0.076$  & $0.21$ \\
\hline
$\beta_{1}$            &    $0.0020$    & $0.11e-4$  & $0.12e-4$ \\
\hline
$\beta_{2}$            &     $2.22e-16$   & $2.22e-16$ & $2.22e-16$\\
\hline
$\beta_{3}$            &     $0.92$   &  $0.60$  & $0.61$\\
\hline
$\beta_{4}$            &     $0.34$   &  $0.27$ & $0.27$\\
\hline
K            &     $2.22e-16$   &  $2.22e-16$   & $2.22e-16$ \\
\hline
P            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$ \\
\hline
$t_{0}$            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$ \\
\hline
$\text{Residuals}$      &     $2.80 \ms$    &  $ 2.29 \ms$ & $2.29 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}
